---
title: "Project1"
author: "Satya Shiva Sai Ram Kamma"
date: "2024-02-11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## News Website Traffic and Revenue Analysing Dataset

I'm creating the dataset for the News website, which mostly depends on website traffic and repeated visitors to earn revenue.

From this Dataset, what I'm trying to find

<b>Which time of day across different traffic sources drives the highest number of sessions and conversions?</b>

   <p> 1. Identifying when each traffic source reaches its peak traffic allows us to optimize your website and its resources, such as content scheduling, ad campaign scheduling, and improving the responsiveness and loading speed of the website.
   
2. Understanding how the time of day, the source of traffic, and website performance relate to one another allows us to make decisions based on data that improve user experience, maximize resource allocation</p>

<b>Is there a significant correlation between content category and session duration?</b>

<p>1. We can learn what type of content is the most valuable and engaging to the audience by studying how long users spend on various content categories.

2. This information can direct the work we do in creating content, ensuring that we focus on topics that interest users and hold their attention for a longer period of time.

3. Identifying parts on a website where users quickly lose interest and quit can be improved by studying the relationship between content categories and session time. This enables to enhance the user experience, layout, or quality of the content in specific categories.</p>
    
<b>Are there any specific landing pages that are particularly successful for different traffic sources or campaigns?</b>


##  Find and collect data
The variables included in this dataset
    
    - Time of Day
    - Traffic Source
    - Landing Page
    - Campaign	
    - Device Category	
    - Avg Session Duration
    - Content Category	
    - Total Sessions	
    - Conversion Rate	
    - Total revenue
    
#### Dependent Variables
    - Total Sessions	
    - Conversion Rate	
    - Total revenue
    
#### Independent Variables
    - Time of Day
    - Traffic Source
    - Landing Page
    - Campaign	
    - Device Category	
    - Avg Session Duration
    - Content Category	

[View the CSV file](News_Website_Dataset.csv)

## Data Dictionary
<p>
<b>Total Sessions:</b> Total number of unique sessions on the website for a specific timeframe (e.g., Day, Week, Month).

<b>Conversion Rate:</b> Percentage of visitors who complete a desired action (e.g., Polls, Newsletter subscription).

<b>Total Revenue:</b> The total amount generated throughout the sessions.

<b>Time of Day:</b> Categorical variable with 4 levels (Morning, Afternoon, Evening, Night).

<b>Traffic Source:</b> Categorical variable indicating where visitors originated from (e.g., Organic Search, Search, Referral, Direct, Social).

<b>Landing Page:</b> The first page a visitor viewed on your website.

<b>Campaign:</b> Categorical variable indicating which marketing campaign a visitor originated from.

<b>Device Category:</b> Categorical variable indicating the device used to access the website (e.g., Desktop, Mobile, Tablet).

<b>Average Session Duration:</b> The average time spent by visitors on your website per session (continuous).

<b>Content Category:</b> Categorical variable classifying the content type of the visited page (e.g., Article, Category, About Us, Contact Us, Home Page(/)).
<p>


```{r}
library(readxl)
News_Website_Dataset <- read_excel("News Website Dataset.xlsx")
#View(News_Website_Dataset)

correlation_coefficient <- cor(News_Website_Dataset$Total_revenue, News_Website_Dataset$Avg_Session_Duration)



plot(News_Website_Dataset$Avg_Session_Duration, News_Website_Dataset$Total_revenue,
     xlab = "Average Session Duration", ylab = "Total Revenue",
     main = "Scatter Plot of Total Revenue vs. Avg Session Duration")

abline(lm(News_Website_Dataset$Total_revenue ~ News_Website_Dataset$Avg_Session_Duration), col = "red")

print(paste("Correlation Coefficient between Total Revenue and Avg Session Duration:", correlation_coefficient))

```
```{r}
# correlation  and coefficient B/W Total_revenue and Total Sessions
correlation_coefficient2 <- cor(News_Website_Dataset$Total_revenue, News_Website_Dataset$Total_Sessions)
print(correlation_coefficient2)

plot(News_Website_Dataset$Total_Sessions, News_Website_Dataset$Total_revenue,
     xlab = "Total Sessions", ylab = "Total Revenue",
     main = "Scatter Plot of Total Revenue vs. Total Sessions")

abline(lm(News_Website_Dataset$Total_revenue ~ News_Website_Dataset$Total_Sessions), col = "blue")
```

## Assignment 2

<b>1. Univariate Analysis:</b>
<p><b>Question :</b>  What is the distribution of total revenue?</p>

<p><b>Visualization:</b> Histogram of Total Revenue</p>
```{r}
library(ggplot2)
hist(News_Website_Dataset$Total_revenue, 
     main = "Distribution of Total Revenue",
     xlab = "Total Revenue",
     ylab = "Frequency",
     col = "skyblue",
     border = "black")
```
<p>The histogram shows the distribution of total income. It means that most of the income falls in the lower ranges, and is distributed to the right. Small amount increases are considered excessive.</p>


<b>2.  Bivariate Analysis:</b>
<p><b>Question :</b> Is there a relationship between total revenue and average session duration?</p>

<p><b>Visualization:</b> Scatter plot of Total Revenue and Avg Session Duration</p>

```{r}
library(ggplot2)
ggplot(News_Website_Dataset, aes(x = Avg_Session_Duration, y = Total_revenue)) +
  geom_point(color = "blue") +
  labs(title = "Total Revenue and Avg Session Duration",
       x = "Average Session Duration",
       y = "Total Revenue")


```
<p>The scatter plot suggests a positive correlation between total revenue and average session duration, as higher revenue tends to coincide with longer session durations.</p>

<b>3.  Bivariate Analysis:</b>
<p><b>Question :</b> How does total revenue vary across different traffic sources?</p>

<p><b>Visualization:</b>  Box plot of Total Revenue by Traffic Source</p>

```{r}
library(ggplot2)
ggplot(News_Website_Dataset, aes(x = Traffic_Source, y = Total_revenue, fill = Traffic_Source)) +
  geom_boxplot() +
  labs(title = "Total Revenue by Traffic Source",
       x = "Traffic Source",
       y = "Total Revenue")
```
<p>The box plot shows variations in total revenue across different traffic sources, with some sources having higher median revenues compared to others.</p>

<b>4.  Multivariate Analysis:</b>
<p><b>Question :</b>How does total revenue vary across different device categories and time of day?</p>

<p><b>Visualization:</b>  Line plot of Total Revenue by Time of Day, color by Device Category</p>

```{r}
library(ggplot2)
ggplot(News_Website_Dataset, aes(x = Time_of_Day, y = Total_revenue, color = Device_Category)) +  geom_line(size = 1.5) +
  labs(title = "Total Revenue by Time of Day (Colored by Device Category)",
       x = "Time of Day",
       y = "Total Revenue")
```
<p>The line plot illustrates how total revenue varies across different times of the day, with each line representing a different device category. It helps identify revenue trends based on the time of day and device usage.</p>

## Assignment 3
##### 1. Total Revenue
<p>Analyzing the univariate mean and variance of the "Total Revenue" variable</p>

```{r}
library(ggplot2)
mean_revenue <- mean(News_Website_Dataset$Total_revenue)
variance_revenue <- var(News_Website_Dataset$Total_revenue)

print(paste("Mean for Total Revenue:", mean_revenue))
print(paste("Variance for Total Revenue:", variance_revenue))

#Box Plot for Total Revenue
ggplot(News_Website_Dataset, aes(x = "", y = Total_revenue)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Box Plot of Total Revenue", x = "", y = "Total Revenue")

# Q-Q plot for Total Revenue
qqnorm(News_Website_Dataset$Total_revenue, main = "Q-Q Plot of Total Revenue")
qqline(News_Website_Dataset$Total_revenue)


```
<p>The mean total revenue provides the average amount of revenue generated across all observations in the dataset. It gives a central measure of the revenue distribution. The variance of total revenue indicates the spread or dispersion of revenue values around the mean. A higher variance suggests that revenue values are more spread out from the mean, while a lower variance suggests that revenue values are closer to the mean. This analysis helps us understand the typical revenue amount and the variability in revenue generation.</p>

#### 2. Total Revenue by Device Category
<p> Analyzing the mean and variance of Total Revenue across different Device Categories </p>
```{r}
library(ggplot2)
mean_revenue_device <- aggregate(Total_revenue ~ Device_Category, data = News_Website_Dataset, mean)
variance_revenue_device <- aggregate(Total_revenue ~ Device_Category, data = News_Website_Dataset, var)

print("Mean Total Revenue by Device Category")
print(mean_revenue_device)
print("Variance of Total Revenue by Device Category")
print(variance_revenue_device)

# Bar plot for Mean Total Revenue by Device Category
ggplot(mean_revenue_device, aes(x = Device_Category, y = Total_revenue, fill = Device_Category)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Mean Total Revenue by Device Category", x = "Device Category", y = "Mean Total Revenue") +
  theme_minimal()

# Violin plot for Distribution of Total Revenue by Device Category
ggplot(News_Website_Dataset, aes(x = Device_Category, y = Total_revenue, fill = Device_Category)) +
  geom_violin(trim = FALSE) +
  labs(title = "Distribution of Total Revenue by Device Category", x = "Device Category", y = "Total Revenue") +
  theme_minimal()


# Box plot for Total Revenue by Device Category
ggplot(News_Website_Dataset, aes(x = Device_Category, y = Total_revenue, fill = Device_Category)) +
  geom_boxplot() +
  labs(title = "Total Revenue by Device Category", x = "Device Category", y = "Total Revenue") +
  theme_minimal()



```
<p>This analysis calculates the mean and variance of total revenue for each device category separately. It helps us understand how revenue varies across different device categories. The mean total revenue by device category provides insight into the average revenue generated by users using each type of device. The variance of total revenue by device category shows the variability in revenue generation among users of the same device category. This analysis can help identify which device categories contribute the most to revenue and how consistent revenue generation is across different devices</p>

### Assignment 4
<p>1 Decide how many Principal Components (PCs) you want to keep and why</p>
<p>2 Explain the variate representation each PCs</p>
<p>3 Perform some visualization using PCs.</p>

```{r}
News_Website_Dataset_num <- read_excel("News Website Dataset.xlsx", range = cell_cols("F:J"))
cor(News_Website_Dataset_num[-1])
News_Website_Dataset_num_pca <- prcomp(News_Website_Dataset_num[,-1],scale=TRUE)
News_Website_Dataset_num_pca
summary(News_Website_Dataset_num_pca)


(eigen_News_Website_Dataset <- News_Website_Dataset_num_pca$sdev^2)
names(eigen_News_Website_Dataset) <- paste("PC",1:4,sep="")
eigen_News_Website_Dataset
sumlambdas <- sum(eigen_News_Website_Dataset)
sumlambdas
propvar <- eigen_News_Website_Dataset/sumlambdas
propvar

cumvar_News_Website_Dataset <- cumsum(propvar)
cumvar_News_Website_Dataset
matlambdas <- rbind(eigen_News_Website_Dataset,propvar,cumvar_News_Website_Dataset)
rownames(matlambdas) <- c("Eigenvalues","Prop. variance","Cum. prop. variance")
round(matlambdas,4)
summary(News_Website_Dataset_num_pca)
News_Website_Dataset_num_pca$rotation
print(News_Website_Dataset_num_pca)

News_Website_Dataset_num_pca$x
```

```{r}
library(FactoMineR)

encoded_data <- model.matrix(~Device_Category - 1, data = News_Website_Dataset_num)
numerical_data <- cbind(News_Website_Dataset_num[, -which(names(News_Website_Dataset_num) == "Device_Category")], encoded_data)

#PCA
pca_result <- prcomp(numerical_data, scale = TRUE)

# Scree plot
plot(pca_result$sdev^2, type = "b", xlab = "Principal Component", ylab = "Variance Explained")

loadings <- pca_result$rotation
print(loadings)

# Data projection onto first two PCs
data_projection <- as.data.frame(pca_result$x[, 1:2])

# Data projection onto next two PCs
data_projection2 <- as.data.frame(pca_result$x[, 3:4])

# Plot
plot(data_projection$PC1, data_projection$PC2, 
     xlab = "PC1", ylab = "PC2", 
     main = "Data Visualization using PCs - 1")

# Plot
plot(data_projection2$PC3, data_projection2$PC4, col = "red",
     xlab = "PC3", ylab = "PC4", 
     main = "Data Visualization using PCs - 2")
```

### Assignment 5
<p>For each model, decide the optimal number of clusters and explain why</p>
<p>Show the membership for each cluster </p>
<p>show a visualization of the cluster and membership using the first two Principal Components</p>
```{r}
# Load necessary libraries
library(cluster)
library(factoextra)
library(magrittr)
library(NbClust)

data <- read_excel("News Website Dataset.xlsx")

data_num <- data[, c("Avg_Session_Duration", "Total_Sessions", "Total_revenue")]
dist_matrix <- dist(data_num)

# Hierarchical clustering
hclust_model <- hclust(dist_matrix)

plot(hclust_model)
num_clusters <- 3
clusters <- cutree(hclust_model, k = num_clusters)

# Membership for each cluster
table(clusters)

# Visualize cluster and membership using first two Principal Components
pca_result <- prcomp(data_num, scale = TRUE)
fviz_cluster(list(data = pca_result$x[, 1:2], cluster = clusters))

# Non-hierarchical clustering (k-means)
num_clusters <- 2  
kmeans_model <- kmeans(data_num, centers = num_clusters)

# Membership for each cluster
table(kmeans_model$cluster)

# Visualize cluster and membership using first two Principal Components
fviz_cluster(list(data = pca_result$x[, 1:2], cluster = kmeans_model$cluster))

```

```{r}
# Load necessary libraries
library(cluster)
library(factoextra)
library(magrittr)
library(NbClust)
library(readxl)

# Read the dataset
data <- read_excel("News Website Dataset.xlsx")

# Select numerical variables for clustering
data_num <- data[, c("Avg_Session_Duration", "Total_Sessions", "Total_revenue")]

# Perform hierarchical clustering
dist_matrix <- dist(data_num)
hclust_model <- hclust(dist_matrix)

# Decide on the optimal number of clusters based on the dendrogram
num_clusters_hclust <- 3  # Replace with chosen number of clusters

# Perform non-hierarchical clustering (k-means)
num_clusters_kmeans <- 2  # Replace with the chosen number of clusters
kmeans_model <- kmeans(data_num, centers = num_clusters_kmeans)

# Visualize cluster centers for k-means
fviz_cluster(kmeans_model, data = data_num, geom = "point", frame.type = "convex", 
             pointsize = 2, fill = "white", main = "K-means Cluster Centers")

# Visualize cluster and membership using first two Principal Components for k-means
pca_result <- prcomp(data_num, scale = TRUE)
fviz_cluster(kmeans_model, data = pca_result$x[, 1:2], geom = "point", 
             pointsize = 2, fill = "white", main = "K-means Clustering Result (PCA)")

# Calculate silhouette information for k-means clustering
sil <- silhouette(kmeans_model$cluster, dist(data_num))

# Visualize the silhouette plot for k-means clustering
fviz_silhouette(sil, main = "Silhouette Plot for K-means Clustering")

# Create a data frame with cluster membership
data_clustered <- cbind(data_num, Cluster = kmeans_model$cluster)

# Scatter plot of data points colored by cluster membership
plot(data_clustered$Avg_Session_Duration, data_clustered$Total_Sessions, 
     col = data_clustered$Cluster, pch = 16, 
     xlab = "Avg_Session_Duration", ylab = "Total_Sessions",
     main = "Scatter Plot of Clustering")
legend("topright", legend = unique(data_clustered$Cluster), 
       col = 1:num_clusters_kmeans, pch = 16, title = "Cluster")

```

### Assignment 6
<h4>Factor Analysis</h4>

<p><b>1. Decide how many Factors are ideal for your dataset</b></p>
<p>Parallel analysis suggests that the number of factors =  1  and the number of components =  1</p>

<p><b>2. Explain the output for your factor model</b></p>
<p><b>Standardized Loadings</b></p>
<p>Avg_Session_Duration has low loadings on all factors, with the highest loading on MR3 (0.32).</p>
<p>Total_Sessions has the highest loading on MR2 (0.34).</p>
<p>Total_revenue and Conversion_Rate both have the highest loadings on MR1 (0.48 and 0.50, respectively).</p>

<p><b>SS Loadings</b></p>
<p>MR1 explains 12% of the total variance., 
MR2 explains 4% of the total variance., 
MR3 explains 3% of the total variance.</p>

<p>MR1 might be related to conversion rate and potentially session duration (positive loadings).</p>
<p>MR2 is positively associated with the total number of sessions.</p>
<p>MR3 has weaker and mixed relationships with the variables.</p>

<p>we can conclude that a 3-factor solution adequately explains the structure of the data. Each factor captures a unique aspect of the underlying structure, with Total_revenue and Conversion_Rate loading primarily on MR1, Total_Sessions on MR2, and Avg_Session_Duration on MR3.</p>

<p><b> 3&4 Show the columns that go into each factor and Perform some visualizations using the factors</b></p>
<p>we can see the below code to answer 3&4.
```{r}
library(psych)
library(readxl)
data <- read_excel("News Website Dataset_2.xlsx")
data_num <- data[, c("Avg_Session_Duration", "Total_Sessions", "Total_revenue","Conversion_Rate")]
factor_model <- fa(data_num, nfactors = 3, rotate = "varimax")

fa.parallel(data_num[-1])

print(factor_model)

factor_loadings <- factor_model$loadings
print(factor_loadings)

#some visualizations
fa.plot(factor_model)      # See Correlations within Factors
fa.diagram(factor_model)   # Visualize the relationship
```
